Casual Bandits

### 2016-01-17 causal-bandit/icml-2015/setup.tex
More on setup

We now introduce a novel class of stochastic sequential decision problems which we call \emph{causal bandit problems}. In these problems, rewards are given for repeated interventions on a fixed causal model~\cite{Pearl2000}. 
The \emph{causal model} is specified using a directed acyclic graph over a set of random variables $\mathcal{X}$ that defines a Bayesian network over $\mathcal{X}$ where an edge from variable $X$ to $X'$ is interpreted to mean that a change in the value of $X$ directly causes a change to the value of $X'$. 
An \emph{intervention (of size $n$)}, denoted $do(\vec{X}=\vec{x})$, assigns the values $\vec{x}=\{x_1, \ldots, x_n\}$ to the corresponding variables $\vec{X}=\{X_1, \ldots, X_n\}$ with the empty intervention (where no variable is set) denoted $do()$.
An intervention ``mutilates'' the original graph by removing edges into the variables in $\vec{X}$ and the resulting graph defines a probability distribution $\P{\vec{X}^c | do(\vec{X}=\vec{x})}$ over $\vec{X}^c := \mathcal{X} - \vec{X}$. 
Further details can be found in Chapter 21 of~\cite{Koller2009}.

A learner for a casual bandit problem is given the casual model, a set of \emph{observable} variables $\mathcal{O} \subseteq \mathcal{X}$, a set of \emph{settable} variables $\mathcal{S} \subset \mathcal{X}$, a \emph{reward variable} $Y \in \mathcal{O} \cap \mathcal{S}^c$, and the following game is played over $T$ rounds. In round $t$, the learner selects an intervention $do(\vec{X}_t = \vec{x}_t)$ for $\vec{X}_t \subseteq \mathcal{S}$. Values for all non-intervened variables $\vec{X}^c_t$ are sampled from $\P{\vec{X}^c_t | do(\vec{X}_t = \vec{x}_t)}$ and the learner is shown the sampled values of all the observable variables in $\vec{X}^c_t \cap \mathcal{O}$, and recevies the reward $Y_t$.
We note that classical $K$-armed stochastic bandit problem can be recovered in our framework by considering a causal model with an edge connecting a single settable variable $X$ that can take on $K$ values to a reward variable $Y = r(X) + \eta$ for some arbitrary but unknown, real-valued function $r$ and a noise term $\eta$ that may depend on $X$.

The expected reward for playing $do(\vec{X}_t = \vec{x}_t)$ is $\mu_{\vec{x}_t} := \E{Y_t | do(\vec{X}_t = \vec{x}_t)}$.

The \emph{cumulative regret} measures the difference between the reward expected under
the omnipotent strategy that knows the optimal action in advance and the expected reward of the learner and
is denoted
\eqn{
\label{eq:regret}
R_T = T \mu^* - \E{\sum_{t=1}^T Y_t}\,.
}
The second performance measure is the \emph{simple regret}, which measures the performance of the learner on the final round only.
\eqn{
\label{eq:regret-simple}
\simpleregret = \mu^* - \E{\mu_{I_T,J_T}}\,.
}
The simple regret typically makes more sense when the learning agent has a fixed learning budget after which its policy will be fixed indefintely. The cumulative
regret is more appropriate for online agents that may continue to adapt their policy.

>>>>

We now introduce a novel class of stochastic sequential decision problems which we call \emph{causal bandit problems}. In these problems, rewards are given for repeated interventions on a fixed causal model~\cite{Pearl2000}. 
The \emph{causal model} is given by a directed acyclic graph $\mathcal{G}$ over a set of random variables $\mathcal{X}$ and a joint distribution $\mathrm{P}$ over $\mathcal{X}$ that factorises over $\mathcal{G}$. An edge from variable $X$ to $X'$ is interpreted to mean that a change in the value of $X$ directly causes a change to the value of $X'$. 
An \emph{intervention (of size $n$)}, denoted $do(\vec{X}=\vec{x})$, assigns the values $\vec{x}=\{x_1, \ldots, x_n\}$ to the corresponding variables $\vec{X}=\{X_1, \ldots, X_n\}$ with the empty intervention (where no variable is set) denoted $do()$.
An intervention ``mutilates'' the original graph by removing edges into the variables in $\vec{X}$ and the resulting graph defines a probability distribution $\P{\vec{X}^c | do(\vec{X}=\vec{x})}$ over $\vec{X}^c := \mathcal{X} - \vec{X}$. 
Further details can be found in Chapter 21 of~\cite{Koller2009}.

A learner for a casual bandit problem is given the casual model's graph $\mathcal{G}$, a set of \emph{observable} variables $\mathcal{O} \subseteq \mathcal{X}$, a set of \emph{settable} variables $\mathcal{S} \subset \mathcal{X}$, a \emph{reward variable} $Y \in \mathcal{O} \cap \mathcal{S}^c$, and the following game is played over $T$ rounds. 
In round $t$, the learner selects an intervention $do(\vec{X}_t = \vec{x}_t)$ for $\vec{X}_t \subseteq \mathcal{S}$. 
Values for all non-intervened variables $\vec{X}^c_t$ are sampled from $\P{\vec{X}^c_t | do(\vec{X}_t = \vec{x}_t)}$ and the learner is shown the sampled values of all the observable variables in $\vec{X}^c_t \cap \mathcal{O}$, and recevies the reward $Y_t$.

We will consider and analyse two different goals for the learner: 1) identifying the intervention with highest expected reward by round $T$, and 2) maximising the total expected reward over all $T$ rounds. 
We denote the expected reward for the intervention $do(\vec{X} = \vec{x})$ by $\mu_{\vec{x}} := \E{Y | do(\vec{X} = \vec{x})}$ and the optimal expected reward by $\mu^* := \max_{\vec{x}} \mu_{\vec{x}}$.
Performance for the first goal can be measured by \emph{simple regret}, \ie, the gap between the optimal expected reward and the intervention chosen by the learner on round $T$:
\eqn{
\label{eq:regret-simple}
\simpleregret = \mu^* - \E{\mu_{\vec{x}_T}}\,.
}
Performance for the second goal is measured by \emph{cumulative regret}, the difference between the total reward obtained if the optimal intervention were known in advance and the expected reward of the learner:
\eqn{
\label{eq:regret}
R_T = T \mu^* - \E{\sum_{t=1}^T Y_t}\,.
}
Simple regret typically makes more sense when the learning agent has a fixed learning budget after which its policy will be fixed indefintely. Cumulative
regret is more appropriate for online agents that may continue to adapt their policy.

We note that classical $K$-armed stochastic bandit problem can be recovered in our framework by considering a simple causal model with one edge connecting a single settable variable $X$ that can take on $K$ values to a reward variable $Y = r(X) + \eta$ for some arbitrary but unknown, real-valued function $r$ and a noise term $\eta$ that may depend on $X$.
Furthermore, any causal bandit problem can reduced to a classical stochastic bandit problem by treating each possible intervention as an arm and ignoring all sampled values for the observed variables except for the reward.
Intuitively though, one would expect
